### 2020-12-03 computational model/TE_computation.Rmd
computational model

---
title: "TE_computation"
date: "Last update: Dec 2nd 2020"
output: html_document
---

# Preparations
## load libraries
```{r setup, include=FALSE}

library(here)
library(tidyverse)
library(tidylog)
library(regclass)

`%notin%` <- Negate(`%in%`)
```

# Generating dataframe for computational model
```{r}
#### Set 1: Calculate other variables based on V, W, and B ####

## Creating variables: V, W, and B
# Items on vocabulary measure (e.g. words on the CDI)
V <- 600 # This is constant

# Total word vocabulary of a child
W <- seq(0, 2*V, by = 50) %>% # a sequence of 0 to 2V
  rep(., 6) # generate 6 repetitinos of the sequence

# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
B <- seq(0.5, 1, by = 0.1) %>% # a sequence of 0.5 to 1
  rep(., 25) # generate 25 repetitions of the sequence

## Creating the dataframe
TE_fromVWB <- data.frame(V, W) %>%
  arrange(W) %>%
  mutate(B = B) %>%
  # Create other variables
  mutate(D = B*W, # Words produced in the dominant language
         N = (1-B)*W, # Words produced in non-dominant language
         TE = (D*N)/V, # If D and N are independent, TE = DN/V
         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
         Du = D-TE, # Unique words in dominant language
         Nu = N-TE) %>% # Unique words in non-dominant language
  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
  filter(Nu >= 0) %>%
  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
  filter_at(vars(D, N), all_vars(. <= 600))

#### Set 2: Calculate other variables based on V, D, and N ####

## Creating variables: D and N
D <- seq(0, V, by = 50) %>% # Words produced in the dominant language
  rep(., 13) # generate 13 repetitions of the sequence

N <- seq(0, V, by =50) %>% # Words produced in non-dominant language
  rep(., 13) # generate 13 repetitions of the sequence

## Creating the dataframe
TE_fromVDN <- data.frame(V, D) %>%
  arrange(D) %>%
  mutate(N = N) %>%
  # It is not possible for N > D, so keep only cases where D >= N
  filter(D >= N) %>%
  # Create other variables
  mutate(W = D+N, # Total word vocabulary of a child
         B = D/(D+N), # Balance, i.e. proportion use of non-dominant language (B ≤ .50)
         TE = (D*N)/V, # If D and N are independent, TE = DN/V
         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
         Du = D-TE, # Unique words in dominant language
         Nu = N-TE) %>%
  # Remove cases where B = NA
  filter(!is.na(B))

#### Combine the 2 sets ####
TE_computation <- TE_fromVWB %>%
  bind_rows(TE_fromVDN)

```

# Correlation
```{r}
Correlation <- TE_computation %>%
  select(-V) %>%
  cor_matrix()

#write.table(Correlation, "TEcomputation_correlations.tsv", col.names=NA)
```

# Plotting relations between variables
## Stacked barchart showing the vocabulary composition across different balance proportions
```{r}

#plot_composition <- TE_computation %>%
#  # Convert to long data
#  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
#  filter(vocab_type %notin% c("D", "N", "C")) %>%
#  mutate(vocab_type = factor(vocab_type, levels=c("TE", "Du", "Nu"))) %>%
#  mutate(bin_W = cut(W, breaks = 3, 
#                     labels = c("0-400", "401-800", "801-1200"))) %>%
#  group_by(bin_W, vocab_type, B) %>%
#  summarise(n_words = mean(number)) %>%
#  ggplot(aes(x=B, y = n_words, fill = vocab_type)) + 
#  geom_bar(position="stack", stat="identity") +
#  facet_grid(. ~ bin_W) +
#  #scale_y_continuous(limits = c(0, 1200)) +
#  theme_minimal() +
#  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") + 
#  labs(x = "Balance", 
#       y = "Number of words")

plot_composition <- TE_computation %>%
  # Convert to long data
  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %notin% c("D", "N", "C")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("Nu", "Du", "TE"))) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  group_by(bin_B, vocab_type, W) %>%
  summarise(n_words = mean(number)) %>%
  ggplot(aes(x=W, y = n_words, fill = vocab_type)) + 
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ bin_B) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
  #scale_y_continuous(limits = c(0, 1200)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") + 
  labs(x = "Total vocabulary size", 
       y = "Number of words", 
       title = "Vocabulary composition across different balance proportions")

ggsave("plot_composition.png", plot_composition)
```

## Plotting relations between TEs and different variables
##### Relation between TE & W, D, & N #####
```{r}
### Plot: Y = TE, X = D / N
plot_TEandWDN <- TE_computation %>%
  pivot_longer(c(W, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %in% c("W", "D", "N")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
  #mutate(bin_B = cut(B, breaks = 6, 
  #                   labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = number, y = TE)) +
  facet_grid(. ~vocab_type) +
  labs(x = "Words produced", y = "Total number of TEs") +
  geom_point(shape = 23) +
  geom_smooth(method="lm", se=F) +
  scale_y_continuous(limits = c(0, 600)) +
  theme_minimal()

ggsave("plot_TEandWDN.png", plot_TEandWDN)

### Plot: Y = TE, X = D / N, color by B
plot_TEandWDN_byB <- TE_computation %>%
  pivot_longer(c(W, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %in% c("W", "D", "N")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = number, y = TE, colour = bin_B)) +
  facet_grid(. ~vocab_type) +
  labs(x = "Words produced", y = "Total number of TEs",
       colour = "Bins of balance") +
  geom_point(shape = 23) +
  geom_smooth(method="lm", se=F) +
  scale_x_continuous(limits = c(0, 1200)) +
  scale_y_continuous(limits = c(0, 600)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") +
  guides(colour = guide_legend(nrow = 1))

ggsave("plot_TEandWDN_byB.png", plot_TEandWDN_byB)
```

##### Relation between W & TE, Du, & Nu #####
```{r}
### Plot: Y = W, X = TE/ Du / Nu
plot_WandTEDuNu <- TE_computation %>%
  pivot_longer(c(TE, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %in% c("TE", "Du", "Nu")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("TE", "Du", "Nu"))) %>%
  #mutate(bin_B = cut(B, breaks = 6, 
  #                   labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = number, y = W)) +
  facet_grid(. ~vocab_type) +
  labs(x = "Words produced", y = "Total vocabulary size") +
  geom_point(shape = 23) +
  geom_smooth(method="lm", se=F) +
  scale_y_continuous(limits = c(0, 1200)) +
  theme_minimal()

ggsave("plot_WandTEDuNu.png", plot_WandTEDuNu)

### Plot: Y = W, X = TE/ Du / Nu, color by B
plot_WandTEDuNu_byB <- TE_computation %>%
  pivot_longer(c(TE, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %in% c("TE", "Du", "Nu")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("TE", "Du", "Nu"))) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = number, y = W, colour = bin_B)) +
  facet_grid(. ~vocab_type) +
  labs(x = "Words produced", y = "Total vocabulary size",
       colour = "Bins of balance") +
  geom_point(shape = 23) +
  geom_smooth(method="lm", se=F) +
  scale_y_continuous(limits = c(0, 1200)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") +
  guides(colour = guide_legend(nrow = 1))

ggsave("plot_WandTEDuNu_byB.png", plot_WandTEDuNu_byB)
```

##### Relation between C & W, D, & N #####
```{r}
### Plot: Y = C, X = W / D / N
plot_CandWDN <- TE_computation %>%
  pivot_longer(c(TE, W, D, N, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %in% c("W", "D", "N")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
  #mutate(bin_B = cut(B, breaks = 6, 
  #                   labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = number, y = C)) +
  facet_grid(. ~vocab_type) +
  labs(x = "Words produced", y = "Number of concepts") +
  geom_point(shape = 23) +
  geom_smooth(method="lm", se=F) +
  scale_y_continuous(limits = c(0, 600)) +
  theme_minimal()

ggsave("plot_CandWDN.png", plot_CandWDN)

### Plot: Y = C, X = W / D / N, color by B
plot_CandWDN_byB <- TE_computation %>%
  pivot_longer(c(TE, W, D, N, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %in% c("W", "D", "N")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = number, y = C, colour = bin_B)) +
  facet_grid(. ~vocab_type) +
  labs(x = "Words produced", y = "Number of concepts",
       colour = "Bins of balance") +
  geom_point(shape = 23) +
  geom_smooth(method="lm", se=F) +
  scale_x_continuous(limits = c(0, 1200)) +
  scale_y_continuous(limits = c(0, 600)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") +
  guides(colour = guide_legend(nrow = 1))

ggsave("plot_CandWDN_byB.png", plot_CandWDN_byB)
```