### 2021-06-09 simulations_old_versions/simulation/vocab_simulation.Rmd

---
title: "Simulation of TE learning"
output: html_notebook
---

This is a first attempt to simulate TE learning in bilinguals.

Hypothetical baby learing 70% English and 30% French

```{r}
library(tidyverse)
library(purrr)
library(corrr)

set.seed(6)

```

Function for simple vocabulary simulation. All words of equal difficulty.

```{r vocab-sim}
## Function that simulates vocab for one baby

# prop_dom is proportion exposure to dominant language
# learn_cutoff is the number of experiences (encounters) needed to learn a word
# experience is the number of times a word could be encountered
# te_cutoff is multiplier for threshold when a TE is lexicalized. te_cutoff = 1 means same experienc needed to learn TE as unique word. Higher values --> harder to learn TEs, i.e. te_cutoff = .5 implies half the experience is needed to learn the TE

VOCAB_SIZE = 100 # Vocab size set here for all simulations.

vocab_sim <- function(prop_dom, learn_cutoff = 30, experience = 50, te= 1) {

  learn_cutoff_first = learn_cutoff # cutoff if first word learned
  learn_cutoff_second = learn_cutoff*te # cutoff if second word learned

  prop_dom = prop_dom # proportion exposure to dominant language
  prop_nondom = 1- prop_dom # proportion exposure to non-dominant language

# Samples how often individual words will be encountered, proportional to exposure
# sample() takes a sample of the specified size from the elements of x using either with or without replacement.

#experience_dom <- prop_dom * sample(1:experience, experience, replace = TRUE)
#experience_nondom <- prop_nondom * sample(1:experience, experience, replace = TRUE)

experience_dom <- prop_dom * sample(1:experience, VOCAB_SIZE, replace = TRUE)
experience_nondom <- prop_nondom * sample(1:experience, VOCAB_SIZE, replace = TRUE)

# Based on above vectors, determine whether each concept is labeled uniquely, or has a TE
vocab <- data.frame(prop_dom, experience_dom, experience_nondom) %>%
  mutate(prop_dom = prop_dom) %>%
  mutate(learn_cutoff = learn_cutoff) %>%
  mutate(experience = experience) %>%
  mutate(te = te) %>%
  mutate(learned_dom = case_when(
    experience_dom >= experience_nondom & experience_dom >= learn_cutoff_first ~ T,
    experience_dom >= experience_nondom & experience_dom < learn_cutoff_first ~ F,
    experience_dom < experience_nondom & experience_dom >= learn_cutoff_second ~ T,
    experience_dom < experience_nondom & experience_dom < learn_cutoff_second ~ F

  )) %>%
  mutate(learned_nondom = case_when(
    experience_nondom >= experience_dom & experience_nondom >= learn_cutoff_first ~ T,
    experience_nondom >= experience_dom & experience_nondom < learn_cutoff_first ~ F,
    experience_nondom < experience_dom & experience_nondom >= learn_cutoff_second ~ T,
    experience_nondom < experience_dom & experience_nondom < learn_cutoff_second ~ F
  )) %>%
#  mutate(learned_dom = if_else(experience_dom >= learn_cutoff, T, F)) %>%
#  mutate(learned_nondom = if_else(experience_nondom >= learn_cutoff, T, F)) %>%
  mutate(learned_unique_dom = if_else(learned_dom & !learned_nondom, T, F)) %>%
  mutate(learned_unique_nondom = if_else(learned_nondom & !learned_dom, T, F)) %>%
  mutate(learned_unique_all = as.logical(learned_unique_dom + learned_unique_nondom)) %>%
  mutate(learned_te = if_else(learned_dom & learned_nondom, T, F)) %>%
  mutate(learned_concepts = if_else(learned_dom | learned_nondom, T, F)) %>%
  select(-experience_dom, -experience_nondom)

return(vocab)
}

# Check that the above function works

vocab_sim(.5, 20, 100, 1.0)

```

Creates summary of simulation

```{r vocab-sim-summary}
## Function that summarizes vocab

vocab_sim_summary <- function(prop_dom, learn_cutoff=15, experience=100, te = 1) {

  vocab <- vocab_sim(prop_dom, learn_cutoff, experience, te) %>%
    group_by(prop_dom, learn_cutoff, experience, te) %>%
  summarize_all(sum)

  return(vocab)

}

## Check that the above function words

vocab_sim_summary(.5, 20, 100)

```

```{r parameter-setup}

# This chunk creates the dataframes/vectors of the space of parameters for the simulation to model

# Create vector with proportions from .1 to .5
seq_prop <-seq(from = .5, to = 1, by = .1)
props <- rep(seq_prop, 25) # 20 repetitions of the different proportion exposures

# For now keep cutoff constant.  Can use this code in the future if necessary

# seq_cutoff <-seq(from = 30, to = 50, by = 10)
# cutoffs <- rep(seq_cutoff, length(props)/length(seq_cutoff))

cutoffs <- rep(50, length(props))

# Create vector with different levels of experience

seq_experience <- seq(from = 50, to = 450, by = 100)
experience <- rep(seq_experience, length(props)/length(seq_experience))

# Create vector for different TE weights.  For now, just use .5, 1, 1.5

seq_te <- c(.5, 1, 1.5)
te <- rep(seq_te, length(props)/length(seq_te))

# Check that the lengths for each vector are equal. These should all return "TRUE"

length(cutoffs) == length(props)
length(experience) == length(props)
length(te) == length(props)

# Get all combinations, then extract columns as lists.  Couldn't find more elegant solution
combo1 <- crossing(props, cutoffs, experience, te) # Full factorial combination of each vector

# Repeat each combo 20 times

combo <- map_dfr(seq_len(20), ~combo1)

# Now need to extract each of the columns separately.  Could be coded more elegantly, but this works for now.

props1 <- combo$props
cutoffs1 <- combo$cutoffs
experience1 <- combo$experience
te1 <- combo$te

# Check how many infants will be simulated

length(props1)

```
Run simulation across all the parameter combinations created in the previous step

```{r}
# Run simulation. This step might take a while

vocab_simulations_full <- 
  pmap(list(props1, cutoffs1, experience1, te1), vocab_sim_summary) %>%
  bind_rows()

# Cleanup and recode some of the columns for easier visualization/interpretation

vocab_simulations <- vocab_simulations_full%>%
  mutate(te = as.factor(as.character(te))) %>%
  mutate(te = recode(te, "0.5" = "te_easier" , "1" = "te_equal", "1.5" = "te_harder")) %>%
  mutate(learn_cutoff = as.factor(as.character(learn_cutoff))) %>%
  mutate(learn_cutoff = recode(learn_cutoff, "30" = "words_easy", "40" = "words_medium", "50" = "words_hard")) %>%
  mutate(learned_words = learned_dom + learned_nondom) %>%
  mutate(prop_dom = as.factor(as.character(prop_dom)))

# Pivot the data into long format

vocab_simulations_long <- vocab_simulations %>%
  pivot_longer(cols = starts_with("learned"), names_to = "word_type", values_to = "n_words")

```

```{r correlations}

# Calculating correlations between different vocabulary measures

all_correl <- vocab_simulations %>%
  ungroup() %>%
    mutate(prop_dom = as.numeric(as.character(prop_dom))) %>%
  filter(prop_dom <= .75) %>%
  mutate(experience = as.numeric(as.character(experience))) %>%
  mutate(learned_propte = learned_te/learned_words) %>%
  relocate(learned_te, learned_propte, learned_dom, learned_nondom, learned_words, learned_concepts) %>%
  group_by(te) %>%
  select(starts_with("learned")| starts_with("prop")|starts_with("experience")) %>%
  group_map(~correlate(.x))

te_neutral <- all_correl[[2]] %>%
  mutate(across(is.numeric, round, 2))

# write_tsv(te_neutral, "correlations.tsv")

```

```{r}

# Plot to compare different vocabulary measures under different TE learning strategies, across experience

plot_words <- vocab_simulations_long %>%
  mutate(prop_dom = as.factor(prop_dom)) %>%
  filter(word_type %in% c("learned_words", "learned_concepts", "learned_dom", "learned_nondom", "learned_te")) %>%
  ggplot(aes(x = experience, y = n_words, color = prop_dom)) +
  geom_point() +
  stat_smooth(method = lm, se = F) +
  theme_minimal() + 
  facet_grid(te ~ word_type) +
  ggtitle("Word vocabulary size by dominance under different measures") +
  theme(axis.text.x = element_text(angle = 90))

ggsave("plot_plot_words.png", plot_words)
plot_words

```

Vocabulary composition across different TE scenarios
```{r}
# Plot of simulations

plot_composition <- vocab_simulations_long %>%
 # filter(te == "te_equal") %>%
 # filter(experience == 350) %>% # Fix a level of experience
  filter(word_type %in% c("learned_te", "learned_unique_dom", "learned_unique_nondom")) %>%
  mutate(word_type = fct_relevel(word_type, "learned_unique_nondom", "learned_unique_dom", "learned_te")) %>%
 # group_by(learn_cutoff) %>%
  group_by(prop_dom, word_type, learn_cutoff, experience, te) %>%
  summarize(n_words = mean(n_words)) %>%
 # mutate(word_type = fct_rev(word_type)) %>%
  ggplot(aes(x = prop_dom, y = n_words, fill = word_type)) +
   geom_bar(position = "stack", stat = "identity", reverse = TRUE) +
  facet_grid(te ~ experience) +
  theme_minimal() +
    ggtitle("Vocabulary composition across different TE scenarios, with different proportions of dominant language exposure") +
     theme(axis.text.x = element_text(angle = 90), legend.position = "bottom")

ggsave("plot_composition.png", plot_composition)
plot_composition

```

```{r}

# non-dominant vocab predicts TEs, but stronger prediction for less balanced infants

vocab_simulations %>%
  filter(te == "te_equal") %>%
  mutate(experience = as.factor(as.character(experience))) %>%
  ggplot(aes(x = learned_nondom, y = learned_te, color = prop_dom)) +
 # facet_grid(. ~ experience) +
  geom_point() +
  theme_minimal() +
 # stat_smooth(method = "lm") +
    ggtitle("Predicting TEs from non-dominant vocabulary")

```

```{r}
# Basically shows that simulations converge on the same thing as computations that are now in the paper

vocab_simulations %>%
filter(te == "te_equal") %>%
  mutate(predict_te = learned_dom * learned_nondom / VOCAB_SIZE) %>%
  ggplot(aes(x = learned_te, y = predict_te, color = prop_dom)) +
    geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  theme_minimal()

```