### 2021-12-06 03_prepare_data.Rmd

---
title: "TE_bilingual_vocabulary_model: Prepare the keepers_data for analysis"
author: "Rachel K.Y. Tsui, Ana Maria Gonzalez-Barrero, Esther Schott, & Krista Byers-Heinlein"
date: '`r format(Sys.time(), "%a %b %d %X %Y")`'
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
```

# Intro
This script implements and documents how we prepared the keepers_data for analysis.

First, we have to read in the keepers_data who holds the bilingual data relevant for the current study.

```{r load_data_raw, cache = TRUE}
keepers_ws_TE <- read.csv(here::here("data_keepers/keepers_ws_TE.csv"))
```

# Pivot data from long to wide
```{r cache = TRUE}
keepers_ws_TE_final <- keepers_ws_TE %>%
  # remove irrelevant variables
  select(-c(vocab_type_dom, vocab_cross_lang, n_administrations)) %>% 
  # change data from long to wide format by spreading "vocab_type"
  pivot_wider(names_from = vocab_type, values_from = vocab_score)
```

# Define language of dominance 
The dominant language was deemed to be the language in which the child produced a greater number of words (i.e., that language with a greater vocabulary size).

* If the child produced a greater number of words in English than French, then English is the dominant language.
* If the child produced a greater numnber of words in French than English, then French is the dominant language.

```{r cache = TRUE}
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  # define lang_dom_vocab
  mutate(lang_dom_vocab = case_when(total_words_eng > total_words_fre ~ "English",
                                 TRUE ~ "French")) %>%
  # make lang_dom_vocab a factor with 2 levels
  mutate(lang_dom_vocab = as.factor(lang_dom_vocab)) 
```

# Define vocabulary measures based on language dominance
Here, we re-defined vocabulary measures with reference to the language of dominance defined in the previous step. The vocabulary measures include:

* total number of words in the dominant language
* total number of words in the non-dominant language
* total number of singlets in the dominant language
* total number of singlets in the non-dominant language
* total number of singlets (i.e., across both languages)

```{r cache = TRUE}
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  mutate(
    # define total number of words in the DOMINANT language
    total_words_dom = case_when(total_words_eng > total_words_fre ~ total_words_eng,
                            TRUE ~ total_words_fre),
    # define total number of words in the NON-DOMINANT language
    total_words_nondom = case_when(total_words_eng > total_words_fre ~ total_words_fre,
                               TRUE ~ total_words_eng),
    # define total number of singlets in the DOMINANT language
    total_singlet_dom = case_when(total_words_eng > total_words_fre ~ eng_unique_words,
                                  TRUE ~ fre_unique_words),
    # define total number of singlets in the NON-DOMINANT language
    total_singlet_nondom = case_when(total_words_eng > total_words_fre ~ fre_unique_words,
                                     TRUE ~ eng_unique_words),
    # compute total number of singlets across both languages
    singlet_vocab = total_singlet_dom + total_singlet_nondom)
```

# Define language balance
A balance score is calculated for each child.

* **balance_vocab**: **The main balance measure** we used in the current study is a *Vocabulary Balance*, which is determined based on the proportion of words produced in the non-dominant language relative to the total words produced across both languages. The formula for calculating the vocabulary balance score is NONDOM/(DOM+NONDOM).
* **balance_input**: Balance can also be considered in terms of input in each language. To make balance_vocab and balance_input comparable, the language designated as DOM and NONDOM was based on vocabulary-defined dominance (lang_dom_vocab), rather than the language that children heard most and least often. The formula for calculating the input balance score is also NONDOM/(DOM+NONDOM).

```{r cache = TRUE}
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  # calculate Vocabulary Balance balance_vocab
  mutate(balance_vocab = total_words_nondom/(total_words_dom+total_words_nondom)) %>%
  # calculate Input Balance balance_input
  mutate(
    # define input dominance (with reference to the lang_dom_vocab defined by vocab size)
    lang_dom_input = case_when(lang_dom_vocab == "English" ~ lang_exp_eng,
                              TRUE ~ lang_exp_fre),
    lang_nondom_input = case_when(lang_dom_vocab == "English" ~ lang_exp_fre,
                                 TRUE ~ lang_exp_eng),
    # calculate the balance score
    balance_input = lang_nondom_input/(lang_dom_input+lang_nondom_input))
```

# Append percentile data to the keepers_data
Finally, we add the percentile data obtained from wordbankr to the keepers_data. This percentile data indicates the number of potentially learnable words for a child at the relevant age.

```{r cache = TRUE}
# read in the percentile data
ws_percentile <- read.csv(here::here("data_keepers/ws_percentile.csv"))

# merge the percentile data to the keepers_ws_TE_final data
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  # as the upper age limit of the CDI-WS is 30m, create a new age variable which replaces all >30m as 30m for merging with the percentile dataframe (this means the percentile should be the same for all >30m)
  mutate(age_months_percentile = ifelse(age_months_binned > 30, 30, age_months_binned)) %>%
  # merge with the percentile data frame
  left_join(ws_percentile, by = "age_months_percentile") 
```

# Save the final keepers_data
The final keepers_data is saved in the folder "data_keepers"
```{r cache = TRUE}
write.csv(keepers_ws_TE_final, "data_keepers/keepers_ws_TE_final.csv", row.names = F)
```